Revisiting the Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation using complex Empirical Orthogonal Functions

### 2021-08-13 analysis/paper/shceof.Rmd
Quita colores de tablas

combine_lists <- function(...) {
  # browser(expr = length(x) != length(y))
  X <- list(...)
  names <- names(X[[1]])
  names(names) <- names
  lapply(names, function(n) {
    unlist(lapply(X, function(x) x[[n]]))
  })
}
```

@@

\textcolor{red}{En la figura 1 te quedaron muy chicos los subplots, amplialos un poco}.

@@

(ref:corr-ceof-splitted-cap) Coefficient of determination ($R^2$) of the absolute magnitude of complex EOFs between 200 hPa and 50 hPa computing EOF separatedly for each level. Background colour indicates the value of $R^2$.

\textcolor{red}{Los colores de la Tabla 1 no dicen mucho en la medida que no se conoce cuál sería la paleta. Quizás se pueden remarcar en negrita ambos aclarando que son significativos (lo son?)}.

```{r corr-ceof-splitted}
ceof_splitted[, eof[[1]]$left, by = .(lev = levs)] %>%
  copy() %>%
  .[, mag  := abs(hgt)] %>%
  .[, item := paste(lev, cEOF, sep = "_")] %>%
  widyr::pairwise_cor(item, time, mag) %>%
  setDT() %>%
  .[tstrsplit(item1, "_")[[1]] != tstrsplit(item2, "_")[[1]], ] %>%
  .[tstrsplit(item1, "_")[[1]] != 50] %>%
  .[, item1 := tstrsplit(item1, "_")[[2]]] %>%
  .[, item2 := tstrsplit(item2, "_")[[2]]] %>%
  .[, r2 := round(correlation^2, 2)] %>%
  dcast(item1 ~ item2, value.var = "r2") %>%
  setnames("item1", "200 hPa") %>%
  kbl2(booktabs = TRUE, caption = "(ref:corr-ceof-splitted-cap)") %>%
  column_spec2(2:4, background = spec_color2(.table[[.col]], option = background_palette,
                                             scale_from = c(0, 1)),
               color = ifelse(.table[[.col]] > .5, "white", "black")) %>% 
  add_header_above(c("", "50 hPa" = 3)) %>%
  kable_classic() 
```

@@

(ref:psa-eof2-cap) Correlations between the Real and Imaginary parts of cEOF2 and the PSA1 and PSA2 modes computed as the second and third EOFs of seasonal geopotential height anomalies [following @mo2001] for the 1979 -- 2019 period. 95% confidence intervals in parenthesis. Background colour indicates the value of correlation.

@@

cEOF2[psa_time, on = "time"] %>% 
  .[PC != "SAM"] %>% 
  copy() %>% 
  sep_ReIm() %>% 
  .[, correlate(value, hgt), by = .(PC, part)] %>% 
  dcast(PC ~ part, value.var = "text") %>% 
  kbl2(booktabs = TRUE, caption = "(ref:psa-eof2-cap)") %>%
  column_spec2(2:3, background = spec_color2(make_numeric(.table[[.col]]),
                                             option = background_palette,
                                             scale_from = c(-1, 1)), 
               color = ifelse(abs(make_numeric(.table[[.col]])) > .5, "white", "black")) %>%
  add_header_above(c("", "cEOF2" = 2)) %>%
  kable_classic()

>>>>

combine_lists <- function(...) {
  # browser(expr = length(x) != length(y))
  X <- list(...)
  names <- names(X[[1]])
  names(names) <- names
  lapply(names, function(n) {
    unlist(lapply(X, function(x) x[[n]]))
  })
}

@@

@@

(ref:corr-ceof-splitted-cap) Coefficient of determination ($R^2$) of the absolute magnitude of complex EOFs between 200 hPa and 50 hPa computing EOF separately for each level. p-values lower than 0.01 in bold.

```{r corr-ceof-splitted}
table <- ceof_splitted[, eof[[1]]$left, by = .(lev = levs)] %>%
  copy() %>%
  .[, mag  := abs(hgt)] %>%
  .[, item := paste(lev, cEOF, sep = "_")]

n <- uniqueN(table$time)

table2 <- table %>%
  widyr::pairwise_cor(item, time, mag) |> 
  setDT() %>%
  .[, p.value := correlation_statistics(correlation, n)$p.value] %>% 
  .[tstrsplit(item1, "_")[[1]] != tstrsplit(item2, "_")[[1]], ] %>%
  .[tstrsplit(item1, "_")[[1]] != 50] %>%
  .[, item1 := tstrsplit(item1, "_")[[2]]] %>%
  .[, item2 := tstrsplit(item2, "_")[[2]]] %>%
  .[, r2 := round(correlation^2, 2)]

pval <- table2 %>% 
  dcast(item1 ~ item2, value.var = c("p.value"))

table2 %>% 
  dcast(item1 ~ item2, value.var = c("r2")) %>% 
  setnames("item1", "200 hPa") %>%
  kbl2(booktabs = TRUE, caption = "(ref:corr-ceof-splitted-cap)") %>%
  column_spec2(2:4, bold = pval[[.col]] < 0.01) %>%
  add_header_above(c("", "50 hPa" = 3)) %>%
  kable_classic() 
```

@@

(ref:psa-eof2-cap) Correlations between the Real and Imaginary parts of cEOF2 and the PSA1 and PSA2 modes computed as the second and third EOFs of seasonal geopotential height anomalies [following @mo2001] for the 1979 -- 2019 period. 95% confidence intervals in parenthesis. p-values lower than 0.01 in bold.

@@

table <- cEOF2[psa_time, on = "time"] %>% 
  .[PC != "SAM"] %>% 
  copy() %>% 
  sep_ReIm() %>% 
  .[, correlate(value, hgt), by = .(PC, part)]

p.val <- table %>% 
  dcast(PC ~ part, value.var = "p.value")

table %>% 
  dcast(PC ~ part, value.var = "text") %>% 
  kbl2(booktabs = TRUE, caption = "(ref:psa-eof2-cap)") %>%
  column_spec2(2:3, bold = p.val[[.col]] < 0.01) %>%
  add_header_above(c("", "cEOF2" = 2)) %>%
  kable_classic()
